---
title: "Final Project Report: Forecasting Pollution Using Time Series Analysis"
author: "Leo Lu (boyuanlu@ucsb.edu)"
date: "2025-03-21"
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
---

# Box-Jenkins approach
```{r 2, message=FALSE, warning=FALSE}
library(tidyverse)
df <- read.csv("filtered_pollution_data.csv")
```

## Step 1: Plot the data
```{r 2.2.1}
plot.ts(df$pollution, main="Time Series of Pollution Levels",
        ylab="Pollution Level", xlab="Time")
```

## Step 2: Transformation
```{r 2.1.2}
df$pollution_log <- log(df$pollution + 1)

plot.ts(df$pollution_log, main="Log Transformed Pollution Data",
        ylab="Log Transformed Pollution Level", xlab="Time")
```

## Step 3: Identification
```{r 2.1.3, message=FALSE, warning=FALSE}
library(forecast)
library(tseries)
par(mfrow=c(1,2))

acf(df$pollution_log, main="ACF of Log Transformed Data")
pacf(df$pollution_log, main="PACF of Log Transformed Data")

par(mfrow=c(1,1))

adf.test(df$pollution_log)
```

## Step 4: Estimation of parameters
```{r 2.1.4}
library(astsa)
fit <- sarima(df$pollution_log, p=3, d=0, q=2)
```

## Step 5: Diagnostics
1. There is no obvious patterns and no outliers exceeding 3 standard deviations.
2. The plot of the estimated ACF of the residuals shows nearly no apparent departure from the model assumption, except a few lags.
3. The normal Q-Q plot of the residuals show that the assumption of normality is reasonable, although there are some outliers.
4. The Q-statistic is not statistically significant.

## Step 6: Model selection
```{r 2.1.6}
df_best <- auto.arima(df$pollution_log, seasonal = FALSE, ic="bic")
summary(df_best)
```

ARIMA(3,0,2) is the best selection for this model.

## Additional Step: Forecast
```{r 2.2}
df_forecast <- forecast(df_best, h=12)
df_forecast1 <- exp(df_forecast$mean)
df_forecast1

time_index <- 1:12
plot(time_index, df_forecast1,
     type = "o", col = "blue",
     xlab = "Forecast Step",
     ylab = "Pollution Level",
     main = "12-Step Ahead Forecast of Pollution")
```

# Unit Root
## Step 1: DF Test
```{r}
df_result <- adf.test(df$pollution_log, k = 0)
print(df_result)
```

## Step 2: ADF Test
```{r}
adf_result <- adf.test(df$pollution_log)
print(adf_result)
```

## Step 3: PP Test
```{r}
pp_result <- pp.test(df$pollution_log)
print(pp_result)
```

# ARMAX Method
## Selection criteria
```{r message=FALSE, warning=FALSE}
library(vars)
dfvars <- cbind(pollution_log = df$pollution_log,
                   TEMP = df$temp,
                   DEW = df$dew,
                   PRESS = df$press)

dfvars <- na.omit(dfvars)
VARselect(dfvars, lag.max = 10, type = "both")
dfarmax <- VAR(dfvars, p = 2, type = "both")
summary(dfarmax)
```
```{r}
dfarmax <- VAR(dfvars, p = 2, type = "both")
summary(dfarmax)
```

## Residual ACF/CCF and Prediction
```{r, fig.width=10, fig.height=6}
serial.test(dfarmax, lags.pt = 12, type = "PT.adjusted")

dfpredict <- predict(dfarmax, n.ahead = 24, ci = 0.95)
fanchart(dfpredict)
```
