---
title: "Project Part-2"
author: "Leo Lu"
date: "2024-12-04"
output:
  pdf_document:
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: paged
    highlight: tango
    latex_engine: xelatex
  html_document:
    df_print: paged
    toc: true
  word_document:
    toc: true
fontsize: 12pt
geometry: margin=1in
fontfamily: mathpazo
chunk_output_type: console
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(readr)
library(skimr)
library(faraway)
```

# Brief Overview of the Data
```{r load_data}
# Data Loading
diamonds <- read_csv("Diamonds Prices2022.csv")

# Data Preview
head(diamonds)

# Adjusting Data
diamonds = diamonds[,-1]
str(diamonds)
diamonds$cut = factor(diamonds$cut,
        levels = c('Fair', 'Good', 'Very Good', 'Premium', 'Ideal'))
diamonds$color = factor(diamonds$color,
        levels = c('D', 'E', 'F', 'G', 'H', 'I', 'J'))
diamonds$clarity = factor(diamonds$clarity,
        levels = c('I1', 'SI2', 'SI1', 'VS2', 'VS1', 'VVS2', 'VVS1', 'IF'))

str(diamonds)
```

# Random Sample
```{r random sampling}
set.seed(111)
index = sample(1:nrow(diamonds), size = 500, replace = F) 
diamonds_sample = diamonds[index, ]
```

# Multiple Linear Regression
```{r linear-regression}
fit <- lm(price ~ carat + depth + cut + color + table, data = diamonds_sample)
summary(fit)
```

# Checking Multicolinearity
## Variance Inflation Factor
```{r variance inflation factor}
X = model.matrix(fit)[,-1]
R = cor(X)
diag(solve(R))

W <- scale(X)
W <- W / sqrt(nrow(X) - 1)
diag(solve(t(W)%*%W))

faraway::vif(fit)
```
These three approaches will give us three same answers where we can observed that there are two VIFs bigger than 10. However, both of cutPremium and cutIdeal are dummy variables from categorical variables, so it is not appropriate to exclude that covatiate.

## Using Condition Index
```{r condition index}
eigenval = eigen(R)$values
sqrt(max(eigenval)/eigenval)
```
As we can see for the results that there are no one number bigger than 30 or even bigger than 10, so there is no severe multicolinearity.

# Variable Selection of Regression Model
## Backward Stepwise Elemination
```{r backward stepwise elemination}
library(MASS)
fit1 <- lm(price ~ carat + depth + cut + color + table, data = diamonds)
fit1_step <- stepAIC(fit1, direction="both", k = 2, trace = 0)
summary(fit1_step)
```

## Forward Stepwise Elemination
```{r forward stepwise elemination}
fit2 = lm(price ~ 1, data = diamonds)
fit2_step <- stepAIC(fit2, direction="both",
                     scope=list(upper=formula(fit1)), k = 2, trace = 0)
summary(fit2_step)
```

# Cross Validation
```{r cross validation}
KFoldCV_lm = function(data, response, formula, k,
                      seed = FALSE, seed_num, shuffle = FALSE) {
  if (seed == TRUE) {
    set.seed(seed_num)
  }
  if(shuffle == TRUE) {
    shuffle_index = sample(1:nrow(data), replace = FALSE)
    data = data[shuffle_index,]
    response = response[shuffle_index]
  }
  
  folds <- cut(seq(1,nrow(data)),breaks=k,labels=FALSE)
  mse = numeric()
  for(i in 1:k){
    testIndexes <- which(folds == i, arr.ind=TRUE)
    testData <- data[testIndexes, ]
    trainData <- data[-testIndexes, ]
    fit.train = lm(formula, data = trainData)
    mse[i] = 
      (1/length(testIndexes))*sum((response[testIndexes] - 
                                     predict(fit.train, newdata = testData))^2)
  }
  rmse = sqrt(mse)
  cv_k_mse = sum(mse)/k
  cv_k_rmse = sum(rmse)/k
  return(list(CV_MSE = cv_k_mse, CV_RMSE = cv_k_rmse))
}

KFoldCV_lm(response = diamonds$price,
           formula =price ~ carat + depth + cut + color + table,
           data = diamonds,
           k = 10, shuffle = TRUE, seed = TRUE, seed_num = 1)

KFoldCV_lm(response = diamonds$price,
           formula =price ~ carat + cut + color,
           data = diamonds,
           k = 10, shuffle = TRUE, seed = TRUE, seed_num = 1)
```